---
title: "Exp. 4"
author: "Matan Mazor"
date: "2023-03-13"
output: word_document
---

```{r setup, include = FALSE}

library('groundhog')
groundhog.library(
  c(
    'papaja',
    'reticulate',
    'tidyverse',
    'broom',
    'cowplot',
    'MESS', # for AUCs
    'lsr', # for effect sizes
    'pwr', # for power calculations
    'brms', # for mixed effects modeling
    'BayesFactor',# for Bayesian t test
    'jsonlite', #parsing data from sort_trial
    'afex' #for anova
  ), "2022-12-01"
)
r_refs("r-references.bib")
knitr::opts_chunk$set(echo=F,message=F, warning=F)

```

```{r load_and_format_data, echo=FALSE, cache=TRUE}
E4.raw_df <- read_csv('../experiments/letters4occludedRows/data/jatos_resultfiles_batch1/all_data.csv') %>%
  # filter(frame_index==0)%>%
  mutate(subj_id=PROLIFIC_PID,
         correct = as.numeric(correct),
         RT = as.numeric(RT),
         present=as.numeric(present),
         resp = response==presence_key) 

```


```{r exclude, echo=FALSE, cache=TRUE,include=FALSE}

E4.low_accuracy <- E4.raw_df %>%
  filter(test_part=='test1' | test_part=='test2') %>%
  group_by(subj_id) %>%
  summarise(
    accuracy = mean(correct)) %>%
    filter(accuracy<0.5) %>%
  pull(subj_id)

E4.too_slow <- E4.raw_df %>%
  filter(test_part=='test1' | test_part=='test2') %>%
  group_by(subj_id) %>%
  summarise(
    third_quartile_RT = quantile(RT,0.75)) %>%
    filter(third_quartile_RT>5000) %>%
  pull(subj_id)

E4.too_fast <- E4.raw_df %>%
  filter(test_part=='test1' | test_part=='test2') %>%
  group_by(subj_id) %>%
  summarise(
    first_quartile_RT = quantile(RT,0.25)) %>%
    filter(first_quartile_RT<100) %>%
  pull(subj_id)

E4.to_exclude <- c(
  E4.low_accuracy,
  E4.too_slow,
  E4.too_fast
) %>% unique()

E4.df <- E4.raw_df %>%
  filter(!(subj_id %in% E4.to_exclude));
  
```

```{r descriptives, echo=FALSE, cache=TRUE}


E4.overall_descriptives <- E4.df %>%
    filter(test_part=='test1' | test_part=='test2') %>%
    group_by(subj_id) %>%
    summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            slow_RT = quantile(RT,0.75)>5000,
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)));
```

`r E4.raw_df%>%pull(subj_id)%>%unique()%>%length()` participants completed the task. Mean accuracy in the main block experiment was `r E4.overall_descriptives%>%pull(accuracy)%>%mean()%>%printnum()` (SD=`r E4.overall_descriptives%>%pull(accuracy)%>%sd()%>%printnum()`). The mean median response time was `r printnum(E4.overall_descriptives%>%pull(RT)%>%mean()/1000)` seconds (SD=`r printnum(E4.overall_descriptives%>%pull(RT)%>%sd()/1000)`). No participants were excluded based on our pre-registered accuracy lower bound of 50%, and based on our pre-registered criterion for making hasty responses (RT\<100 ms in more than 25% of the trials). `r E4.too_slow%>%length()` participants were excluding for having resposne times longer than 5 seconds in more than 25% of the trials, leaving `r E4.df$subj_id%>%unique()%>%length()` for the main analysis. 


```{r H1, echo=FALSE, cache=TRUE}

E4.RT_by_resp <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100) %>%
  group_by(subj_id,resp) %>%
  summarise(RT=median(RT))%>%
  spread(resp,RT,sep='')%>%
  mutate(diff=respTRUE-respFALSE)
```

*Hypothesis 1 (PRESENCE/ABSENCE RESPONSE TIME)*: As expected, response times were significantly shorter in decisions about presence compared to absence (`r printnum(E4.RT_by_resp%>%pull(respTRUE)%>%mean()/1000)` vs `r printnum(E4.RT_by_resp%>%pull(respFALSE)%>%mean()/1000)` seconds; `r apa_print(E4.RT_by_resp%>%pull(diff)%>%t.test())$statistic`).


```{r H2, echo=FALSE, cache=TRUE}

E4.RT_by_occlusion_in_presence <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & resp) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(RT=median(RT))%>%
  spread(hide_proportion,RT,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

E4.RT_by_occlusion_in_presence_correct_only <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & resp & correct) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(RT=median(RT))%>%
  spread(hide_proportion,RT,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);
```

*Hypothesis 2 (OCCLUSION RT EFFECT IN PRESENCE)*: Target-present decisions were significantly slower when more of the target was occluded behind black rows (`r printnum(E4.RT_by_occlusion_in_presence_correct_only$hide_proportion0.35%>%mean()/1000)` vs `r printnum(E4.RT_by_occlusion_in_presence_correct_only$hide_proportion0.1%>%mean()/1000)` seconds;  `r apa_print(E4.RT_by_occlusion_in_presence_correct_only%>%pull(diff)%>%t.test())$statistic`).

```{r H3, echo=FALSE, cache=TRUE}

E4.RT_by_occlusion_in_absence <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & !resp) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(RT=median(RT))%>%
  spread(hide_proportion,RT,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

E4.RT_by_occlusion_in_absence_correct_only <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & !resp & correct) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(RT=median(RT))%>%
  spread(hide_proportion,RT,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);
```

*Hypothesis 3 (OCCLUSION RT EFFECT IN ABSENCE)*: Target-absent decision times were unaffected by the occlusion manipulation (`r printnum(E4.RT_by_occlusion_in_absence_correct_only$hide_proportion0.35%>%mean()/1000)` vs `r printnum(E4.RT_by_occlusion_in_absence_correct_only$hide_proportion0.1%>%mean()/1000)` seconds for the 2 and 6 rows conditions, respectively;  `r apa_print(E4.RT_by_occlusion_in_absence_correct_only%>%pull(diff)%>%t.test())$statistic`).


```{r H4, echo=FALSE, cache=TRUE}

E4.RT_by_occlusion_and_response <- merge(
  E4.RT_by_occlusion_in_presence,
  E4.RT_by_occlusion_in_absence,
  by= 'subj_id',
  suffixes = c('presence','absence')) %>%
  mutate(interaction = diffpresence-diffabsence);

E4.RT_by_occlusion_and_response_correct_only <- merge(
  E4.RT_by_occlusion_in_presence_correct_only,
  E4.RT_by_occlusion_in_absence_correct_only,
  by= 'subj_id',
  suffixes = c('presence','absence')) %>%
  mutate(interaction = diffpresence-diffabsence);
```

*Hypothesis 4 (OCCLUSION RESPONSE INTERACTION ON RT)*: The effect of occlusion on response time was stronger in decisions about target presence, compared to decisions about target absence (`r apa_print(E4.RT_by_occlusion_and_response_correct_only%>%pull(interaction)%>%t.test())$statistic`).

```{r quantiles, echo=FALSE, cache=TRUE, fig.cap="Response time quantiles 50, 75, 90 and 95 for correct decisions about presence and absence when 2 or 6 rows are occluded." }


# a function to map the quantiles of target absent search times to
# target present search times.
getQuantiles <- function(pres,abs,probs) {
  present = quantile(pres,probs)
  absent = quantile(abs,probs)
  return(data.frame(probs, present,absent))
}


E4.quantiles <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & correct) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(getQuantiles(
    RT[resp],
    RT[!resp],
    c(0.5,0.75,0.9,0.95)
  )) %>%
  pivot_longer(cols=c('present','absent'), names_to='resp', values_to='RT')

E4.quantile_occlusion_effects <- E4.quantiles %>%
  group_by(subj_id, probs,resp) %>%
  spread(hide_proportion, RT, sep='') %>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35) %>%
  dplyr::select(subj_id,probs,diff)%>%
  pivot_wider(names_from=resp,values_from=diff)

E4.quantile_summary <- E4.quantiles %>% 
  group_by(subj_id) %>%
  mutate(centeredRT = RT-mean(RT))%>%
  group_by(hide_proportion,probs, resp) %>%
  summarise(mean_RT=mean(RT),
            se_RT = se(centeredRT)) %>%
  mutate(resp=factor(resp, levels=c('present','absent')),
         hide_proportion=factor(hide_proportion,levels=c(0.1,0.35)))

E4.quantile_plot <- E4.quantile_summary %>% 
  mutate(occluded_rows = factor(ifelse(hide_proportion==0.1,2,6)))%>%
  ggplot(aes(x=occluded_rows, 
             y=mean_RT, 
             color=resp,
             alpha=factor(probs),
             group=probs)) +
  geom_line(size=1.3)+
  geom_errorbar(aes(ymin=mean_RT-se_RT,ymax=mean_RT+se_RT))+
  scale_color_manual(values=c("#377eb8", "#e41a1c"))+
  scale_alpha_manual(values=c(1,0.8,0.6,0.4))+
  facet_grid(~resp)+
  labs(x='number of hidden rows',
       y='RT (ms)')+
  theme_bw() +
  theme(legend.pos='na')

ggsave('figures/E4Quantiles_separate.png',E4.quantile_plot,width=3.5,height=3.5);

E4.quantile_plot

```

```{r SDT, echo=FALSE, cache=TRUE}

E4.descriptives_by_occlusion<- E4.df %>%
  filter(test_part=='test1' | test_part=='test2') %>%
  mutate(resp = response==presence_key)%>%
  group_by(subj_id,hide_proportion) %>%
  summarise(accuracy=mean(correct),
            resp_bias=mean(resp),
            RT=median(RT),
            hit_rate = (sum(correct & present)+0.5)/(sum(present)+1),
            fa_rate = (sum(!correct & !present)+0.5)/(sum(!present)+1),
            d = qnorm(hit_rate)-qnorm(fa_rate),
            c = -0.5*(qnorm(hit_rate)+qnorm(fa_rate)))

E4.dprime <- E4.descriptives_by_occlusion%>%
  dplyr::select(subj_id,hide_proportion,d)%>%
  spread(hide_proportion, d,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35)

E4.criterion <- E4.descriptives_by_occlusion%>%
  dplyr::select(subj_id,hide_proportion,c)%>%
  spread(hide_proportion, c,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35)

E4.hit_rate <- E4.descriptives_by_occlusion%>%
  dplyr::select(subj_id,hide_proportion,hit_rate)%>%
  spread(hide_proportion, hit_rate,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35)

E4.fa_rate <- E4.descriptives_by_occlusion%>%
  dplyr::select(subj_id,hide_proportion,fa_rate)%>%
  spread(hide_proportion, fa_rate,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35)
```

*Hypothesis 5 (SENSITIVITY)*: Signal detection sensitivity was significantly impacted by pixel occlusion (`r apa_print(E4.dprime$diff%>%t.test())$statistic`), with d'=`r printnum(E4.dprime$hide_proportion0.1%>%mean())` (SD=`r printnum(E4.dprime$hide_proportion0.1%>%sd())`) when two rows were occluded, and d'=`r printnum(E4.dprime$hide_proportion0.35%>%mean())` (SD=`r printnum(E4.dprime$hide_proportion0.35%>%sd())`) when six rows were occluded. This effect was driven by a decrease in hit rate (`r apa_print(E4.hit_rate$diff%>%t.test())$statistic`), and a smaller decrease in the false-positive rate (`r apa_print(E4.fa_rate$diff%>%t.test())$statistic`) when six rows were occluded.

*Hypothesis 6 (CRITERION)*: Signal detection criterion was conservative overall, and more conservative when more of the display was occluded (`r apa_print(E4.criterion$diff%>%t.test())$statistic`), with c=`r printnum(E4.criterion$hide_proportion0.1%>%mean())` (SD=`r printnum(E4.criterion$hide_proportion0.1%>%sd())`) when two rows were occluded, and `r printnum(E4.criterion$hide_proportion0.35%>%mean())` (SD=`r printnum(E4.criterion$hide_proportion0.35%>%sd())`) when six rows were occluded. 

```{r H7, echo=FALSE, cache=TRUE}

E4.confidence_by_resp <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2')) %>%
  group_by(subj_id,resp) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(resp,confidence,sep='')%>%
  mutate(diff=respTRUE-respFALSE)

E4.confidence_by_resp_correct_only <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & correct) %>%
  group_by(subj_id,resp) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(resp,confidence,sep='')%>%
  mutate(diff=respTRUE-respFALSE)
```

*Hypothesis 7 (PRESENCE/ABSENCE CONFIDENCE)*: Confidence was descriptively, but not significantly higher in decisions about presence compared to absence (`r printnum(E4.confidence_by_resp_correct_only%>%pull(respTRUE)%>%mean())` vs `r printnum(E4.confidence_by_resp_correct_only%>%pull(respFALSE)%>%mean())` on a 0-1 scale; `r apa_print(E4.confidence_by_resp_correct_only%>%pull(diff)%>%t.test())$statistic`).

```{r H8, echo=FALSE, cache=TRUE}

E4.confidence_by_occlusion_in_presence <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & resp) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(hide_proportion,confidence,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

E4.confidence_by_occlusion_in_presence_correct_only <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & resp & correct) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(hide_proportion,confidence,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

```

*Hypothesis 8 (OCCLUSION CONFIDENCE EFFECT IN PRESENCE)*: Confidence in decisions about absence was lower when more of the display was occluded (`r printnum(E4.confidence_by_occlusion_in_presence_correct_only$hide_proportion0.35%>%mean())` vs. `r printnum(E4.confidence_by_occlusion_in_presence_correct_only$hide_proportion0.1%>%mean())`; `r apa_print(E4.confidence_by_occlusion_in_presence_correct_only$diff%>%t.test())$statistic`). 


```{r H9, echo=FALSE, cache=TRUE}

E4.confidence_by_occlusion_in_absence <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & !resp) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(hide_proportion,confidence,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

E4.confidence_by_occlusion_in_absence_correct_only <- E4.df %>%
  filter((test_part=='test1' | test_part=='test2') & RT>100 & !resp & correct) %>%
  group_by(subj_id,hide_proportion) %>%
  summarise(confidence=mean(confidence, na.rm=T))%>%
  spread(hide_proportion,confidence,sep='')%>%
  mutate(diff=hide_proportion0.1-hide_proportion0.35);

```

*Hypothesis 9 (OCCLUSION CONFIDENCE EFFECT IN ABSENCE)*: Confidence in decisions about presence was also lower when more of the display was occluded (`r printnum(E4.confidence_by_occlusion_in_absence_correct_only$hide_proportion0.35%>%mean())` vs. `r printnum(E4.confidence_by_occlusion_in_absence_correct_only$hide_proportion0.1%>%mean())`; `r apa_print(E4.confidence_by_occlusion_in_absence_correct_only$diff%>%t.test())$statistic`). 

```{r H10, echo=FALSE, cache=TRUE}

E4.confidence_by_occlusion_and_response <- merge(
  E4.confidence_by_occlusion_in_presence,
  E4.confidence_by_occlusion_in_absence,
  by= 'subj_id',
  suffixes = c('presence','absence')) %>%
  mutate(interaction = diffpresence-diffabsence);

E4.confidence_by_occlusion_and_response_correct_only <- merge(
  E4.confidence_by_occlusion_in_presence_correct_only,
  E4.confidence_by_occlusion_in_absence_correct_only,
  by= 'subj_id',
  suffixes = c('presence','absence')) %>%
  mutate(interaction = diffpresence-diffabsence);
```

*Hypothesis 10 (OCCLUSION RESPONSE INTERACTION ON CONFIDENCE)*: The effect of occlusion on confidence was not significantly different in decisions about target presence and absence (`r apa_print(E4.confidence_by_occlusion_and_response_correct_only%>%pull(interaction)%>%t.test())$statistic`).


```{r confidence-plot, echo=FALSE, cache=TRUE, fig.cap="Mean confidence in correct decisions about presence and absence when 2 or 6 rows are occluded." }


E4.confidence <- E4.df %>%
  mutate(resp=factor(ifelse(resp,'present','absent'), 
                     levels=c('present','absent'))) %>%
  filter((test_part=='test1' | test_part=='test2') & correct) %>%
  group_by(subj_id,hide_proportion,resp) %>%
  summarise(confidence=mean(confidence,na.rm=T)) 

E4.confidence_summary <- E4.confidence %>%
   group_by(hide_proportion,resp) %>%
  summarise(mean_conf=mean(confidence),
            se_conf=se(confidence))

E4.confidence_plot <- E4.confidence %>% 
  mutate(occluded_rows = factor(ifelse(hide_proportion==0.1,2,6)))%>%
  ggplot(aes(x=occluded_rows, 
             y=confidence, 
             color=resp)) +
  geom_jitter(width=0.2,alpha=0.5, size=3)+
  # geom_errorbar(aes(ymin=mean_RT-se_RT,ymax=mean_RT+se_RT))+
  scale_color_manual(values=c("#377eb8", "#e41a1c"))+
  facet_grid(~resp)+
  geom_boxplot(fill='white',alpha=0.5, size=1, outlier.shape=NA)+
  labs(x='number of hidden rows',
       y='confidence')+
  theme_classic() +
  theme(legend.pos='na')

ggsave('figures/E4confidence.png',E4.quantile_plot,width=3.5,height=3.5);

E4.confidence_plot

```


